---
title: "Canada EDA Visualization - Multiple Line Plots"
author: "Group C"
date: "12/21/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Descriptive and Exploratory Question Analysis
#### Q4 - What are the immigrants counts for the top 5 immigrants countries from 1980 - 2013 Visualized by line charts and area plot.

#### Import the necessary libraries

```{r}
library(tidyverse) #For data manipulation and ggplot
library(dplyr) #For Data manipulation
library(ggpubr) #For sub plotting
```

#### Import the Dataset

```{r}
DF <- read.csv('/Users/salahkaf/Desktop/UpdatedDF.csv') #Reading the DF
head(DF)
```

#### Cleaning the dataset

```{r}
DFQ4 <- DF[-c(196,197),] # Removing "unknown" and "Total" rows
```

#### Sorting by descending order to find top 5

```{r}
DFtop5 <- arrange(DFQ4,desc(Total))
DFtop5 <- DFtop5[1:5,c(1,5:38)]
DFtop5
```

#### Manipulate the table to make it suitable for visualization 

```{r}
Q4DF <- t(DFtop5) #Take the transpose of df
Q4DF <- as.data.frame(Q4DF) #Convert the transposed data into data frame
Q4DF <- cbind(Year = rownames(Q4DF), Q4DF) #Make the index as a column
rownames(Q4DF) <- 1:nrow(Q4DF) #Create a new index
Q4DF$Year<-gsub("X","",as.character(Q4DF$Year)) #Removing X from years
Q4DF <- Q4DF[-1,] #Remove the unnecessary column
colnames(Q4DF) <- c("Year","India","China","UK","Philippines","Pakistan") #Rename the columns
#Make all columns as integer data type
for (i in 1:ncol(Q4DF)) {
Q4DF[,i] <- as.integer(Q4DF[,i])
}
Q4DF #Final DF
```

#### Plotting the dataset

```{r}
colors <- c("India", "China", "UK","Philippines","Pakistan")

Q4plot_multiplelines <- ggplot(data = Q4DF, aes(x= Year, group = 1)) + 
  geom_line(aes(y = India, color = "India"),linetype="twodash") + 
  geom_line(aes(y = China,color = "China"),linetype="twodash") +
  geom_line(aes(y= UK, color = "UK"),linetype="twodash") +
  geom_line(aes(y= Philippines, color = "Philippines"),linetype="twodash") +
  geom_line(aes(y= Pakistan, color = "Pakistan"),linetype="twodash") +
  scale_x_continuous(breaks = seq(1980, 2013, by = 5))+ #A jump of 5 years
  scale_y_continuous(breaks = seq(10000, 50000, by = 10000)) +
  ggtitle("Number of Immigrants per Year") +
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x = "Year",y = "Immigrants") 
Q4plot_multiplelines
```

#### Plot Area plot 

```{r}
Q4plotarea <- ggplot(data = Q4DF, aes(x=Year,group = 1)) +
geom_area(aes(y= India, fill='India',alpha= 0.65)) +
geom_area(aes(y= China,fill='China',alpha= 0.65)) +
geom_area(aes(y= UK,fill= 'UK',alpha= 0.65)) +
geom_area(aes(y= Philippines,fill= 'Philippines',alpha= 0.65)) +
geom_area(aes(y= Pakistan,fill= 'Pakistan',alpha= 0.65)) +
scale_x_continuous(breaks = seq(1980, 2013, by = 5))+ #A jump of 5 years
scale_y_continuous(breaks = seq(10000, 50000, by = 10000)) +
ggtitle("Number of Immigrants per Year") +
theme(plot.title = element_text(hjust = 0.5))+
labs(x = "Year",y = "Immigrants") 
Q4plotarea
```

