---
title: 'Canada EDA Visualization - Cloud & Waffle Charts'
author: "Group C"
date: "12/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Exploratory Question Analysis
#### Q7 - Which countries would have a dominant immigrants’ population? Visualized by waffle chart and word cloud.

#### Import the necessary libraries

```{r}
library(tidyr) #For data manipulation and ggplot
library(dplyr) #For data manipulation
library(ggpubr) #For sub plotting
library(waffle) # for plotting waffle chart
library(wordcloud) # for plotting word cloud
```

#### Import the Dataset

```{r}
DF <- read.csv('/Users/salahkaf/Desktop/UpdatedDF.csv') #Reading the DF
head(DF)
```


#### Wrangling the table part 1

```{r}
Q7DF <- DF[c("Country", "Total")] #Choosing these two columns only
Q7DF <- Q7DF[-c(nrow(Q7DF), nrow(Q7DF)-1),] # removing the last two rows ["Unknown" "Total"]

```

#### Plot the word cloud

```{r}
Q7Word_plot <- wordcloud(words = Q7DF$Country, freq = Q7DF$Total, min.freq = 1,
          max.words=2000, random.order=FALSE, rot.per=0.35,
          colors=brewer.pal(8, "Dark2"))
Q7Word_plot
```


#### Wrangling the table part 2

```{r}
Q7DF2 <-arrange(Q7DF, desc(Total))
Q7DF2 <- Q7DF2[1:8,] #Choosing the first 8 rows >> top 8
Q7DF2$Total <- Q7DF2$Total%/%10000 # scaling down the values by a factor of 10000 to be plottable 
Q7DF2
waffle_list <- split(Q7DF2$Total, Q7DF2$Country ) # used split() to make a list that contains countries names and their values which is the total number of immigrants
waffle_list <- unlist(waffle_list) # unlist the data because waffle() doesn't support lists
```

#### Plot the waffle chart

```{r}
waffle(waffle_list)
```

#### selecting individual countries for waffle chart

```{r}
Q7DF3 <- filter(Q7DF, Total >100) 
Q7DF3 <- filter(Q7DF3, Country %in% c("India" ,"China", "Sudan")) # the chosen countries ## To be automated
Q7DF3$Total <- Q7DF3$Total%/%min(Q7DF3$Total)
waffle_list2 <- split(Q7DF3$Total, Q7DF3$Country ) # used split() to make a lsit that contains countries names and their values which is the total number of immigrants
waffle_list2 <- unlist(waffle_list2) # unlist the data because waffle() doesn't support lists
```

#### Plot the waffle chart for three chosen countries

```{r}
Q7waffle <- waffle(waffle_list2)
Q7waffle 
```
