---
title: "Canada EDA Visualization - Pie plot and line plot"
author: "Group C"
date: "12/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Descriptive Analysis Question
#### Q2 - How many total immigrants to Canada by continent from 1980 to 2013. Visualized by pie and line charts.

#### Part 1 - Pie Chart

#### Import the necessary libraries

```{r}
library(tidyverse) #For data manipulation and ggplot
library(dplyr) #For data manipulation
library(ggpubr) #For sub plotting
```

#### Import the Dataset

```{r}
DF <- read.csv('/Users/salahkaf/Desktop/UpdatedDF.csv') #Reading the DF
head(DF)
```

#### Wrangling the table part 1

```{r}
DF <- DF[-c(196,197),] #Remove last two rows as they do not have a continent
Q2DF1 <- tapply(DF$Total,DF$Continent,sum) #Sum total immigrants per continent
Q2DF1 <- as.data.frame.table(Q2DF1) #Convert it to a data frame
Q2DF1
```
# Wrangling the table part 2

```{r}
names <- c("Africa", "Asia", "Europe", "Latin America and the Caribbean","Northern America","Oceania") #Names of the continents
percentage <- round(Q2DF1$Freq/sum(Q2DF1$Freq)*100,2) #Count the percentage of each Continent
lebals <- paste(names, percentage) # add percents to labels 
lebals <- paste(lebals,"%",sep="") # add % to labels 
Q2DF1$Continent <- lebals #Add the labels as a new column "Continent"
Q2DF1<- Q2DF1[,-1] #Remove the old column of names
names(Q2DF1)[1] <- "Total" #Rename Freq to Total
Q2DF1 #Final DF for part 1
```
#### Plot the Pie Chart

```{r}
Q2plot_Pie <- ggplot(Q2DF1, aes(x="", y=Total, fill=Continent)) +
  geom_bar(stat="identity", width=1, color="White") +
  coord_polar("y", start=0) + 
  ggtitle("Percentage of Immigrants per Continent [1980-2013]") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_void() # remove background, grid, numeric labels
Q2plot_Pie
```

#### Line plot of [Total number of immigrants per continant from 1980 to 2013]

#### Creating a suitable DF for the line plot - Part A

```{r}
Continent <- c("Africa","Asia","Europe","Latin America and the Caribbean","Northern America","Oceania")
Q2DF2 <- data.frame(Continent)
Q2DF2
```

```{r}
years <- names(DF[,-c(1,2,3,4,39)]) #Keeping only the years columns

# looping through each column and sum number of immigrants per continent
for (col in years) {
summation <- tapply(DF[,col],DF$Continent,sum) #Sum total immigrants per continent for each year
Q2DF2[col] <- summation
}
Q2DF2
```

#### Creating a suitable DF for the line plot - Part B

```{r}
Q2DF2 <- Q2DF2[Q2DF2$Continent=="Africa",2:35] #Here "Africa" is the user input, the purpose is to automate this option
Q2DF2
```
#### The steps of line plot is repeated as before in Q1
#### Transposing the subsetted dataset

```{r}
Q2DF2 <- gather(Q2DF2,Year,Total,1:34) #Transposing the dataset 
Q2DF2$Year <- gsub("X","",as.character(Q2DF2$Year)) #Removing X from years
Q2DF2 #Final DF ready to be plotted
```

```{r}
#### Plotting Code - Line plot
Q2plot_line <- ggplot(data=Q2DF2, aes(x=Year, y=Total, group=1)) +
  geom_line(color = "Green")+ #Draw a line
  geom_point() + #Draw points 
   scale_x_discrete(breaks = seq(1980, 2013, by = 5))+ #A jump of 5 years
   scale_y_continuous(breaks = seq(10000, 50000, by = 5000)) +
  ggtitle("Number of Immigrants per Year") +
  theme(plot.title = element_text(hjust = 0.5))+
  xlab("Years") + ylab("Number of Immigrants")
Q2plot_line
```

