---
title: "Data Cleaning"
author: "Ahmad Abdalla"
date: "12/18/2021"
output: html_document
---
## Importing the required libraries
```{r}
library("readxl")
library("dplyr")
```
## Reading the data excel file

```{r}
sheet2 <- read_excel("C:/ahmed/UM/University Of Malaya/YEAR 1 # FIRST SEMESTER/Principle of Data Science/Project/Dataset/Canada.xlsx",
           sheet ="Canada by Citizenship" )
```
## Removing unnecessary rows
```{r}
sheet2 <- sheet2[-(1:15),]
names(sheet2) <- as.character(unlist(sheet2[1,]))
sheet2 <- sheet2[-1,]
```
## removing unnecessary coloumns
```{r}
sheet2 <- select(sheet2, OdName, AreaName, RegName,DevName, "1980":"2013")
```
## Converting the data into a data frame format 
```{r}
sheet2 <- as.data.frame(sheet2)
```
## Modifying the names of the columns

```{r}
names(sheet2)[1:4] <- c("Country", "Continent", "Region","DevName" )
```
## Adding a "Total" column to sum the vlues 

```{r}
Total <- apply(sheet2[-(1:4)], 1, sum)
sheet2 <- mutate(sheet2, Total = Total)
```
## adding "X" before each year column

```{r}
colnames(sheet2)[c(-(1:4), -39)] <- paste("X", colnames(sheet2)[c(-(1:4), -39)], sep = "")
```
## Modifying row 44 country name
```{r}
sheet2[44,1] <- "Côte d'Ivoire"

```
## checking if the cleaned dataset maches our new dataset
```{r}
cleaned <- read.csv("C:/ahmed/UM/University Of Malaya/YEAR 1 # FIRST SEMESTER/Principle of Data Science/Project/Dataset/canadian_immegration_data.csv")
setdiff(sheet2, cleaned)
```
## Viewing the data
```{r}
View(sheet2)
```
## exporting the data frame 
```{r}
New_DataFrame <- sheet2
write.csv(New_DataFrame,"C:/ahmed/UM/University Of Malaya/YEAR 1 # FIRST SEMESTER/Principle of Data Science/Project/Dataset/New_DataFrame.csv",row.names = F)

```
