---
title: "Immigration to Canada - Dataset Cleaning"
author: "Group C"
date: "12/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### This dataset presents the annual counts of international immigrants to Canada from 1980 - 2013. It was downloaded from the International migration flows to and from selected countries - The 2015 revision.

#### Step 1 - Import the necessary libraries 

```{r}
library('readxl')
library('dplyr')
```

#### Step 2 - Reading the original dataset

```{r}
OriginalDF <- read_excel("/Users/salahkaf/Desktop/Principle\ Project/Canada.xlsx", sheet ="Canada by Citizenship",skip = 20) # We removed the first 20 rows as they are metadata about the dataset.
```

#### Step 3 - Get some basic understanding about the dataset.

```{r}
head(OriginalDF) #Shows the first 6 rows
tail(OriginalDF) #Shows the last 6 rows
dim(OriginalDF) #Shows the dimensions of it
str(OriginalDF) #Shows the structure of it
summary(OriginalDF) #Shows the summary of it
```
#### Step 4 - Checking for missing values

```{r}
# Total number of missing values in the dataset
cat("The total number of missing values in the dataset is" , sum(is.na(OriginalDF)))

# Total number of missing values in the dataset per column name
colSums(is.na(OriginalDF)) 

# Total number of missing values in the dataset per column index
for (i in 1:ncol(OriginalDF)) {
x <- sum(is.na(OriginalDF[i]))
cat("Number of missing values in col ", i, " = ", x, "\n")
}
```

# Step 5 - Making all columns names as charectar 

```{r}
names(OriginalDF) <- as.character(unlist(names(OriginalDF)))
class(names(OriginalDF)) #Checking the class is character
```

# Step 6 - Selecting the necessary columns for analysis
```{r}
UpdatedDF <- select(OriginalDF, OdName, AreaName, RegName,DevName, "1980":"2013")
head(UpdatedDF)
```
# Step 7 - Modify the name of columns

```{r}
names(UpdatedDF)[1:3] <- c("Country", "Continent", "Region")
head(UpdatedDF)
```
# Step 8 - Add Total column at the end

```{r}
UpdatedDF <- mutate(UpdatedDF, Total=apply(UpdatedDF[-c(1:4)],1,sum))
```

# Step 9 - Export the cleaned dataset
```{r}
write.csv(UpdatedDF,"/Users/salahkaf/Desktop/UpdatedDF.csv",row.names = F)
```



